---
title: "HW1-108020033"
author: "Che-Wei, Chang"
output:
  pdf_document: default
  html_document:
    df_print: paged
date: "2023-02-17"
---
```{r, setup, include=FALSE}
knitr::opts_chunk$set(message = TRUE)
```
### 1.What is the 5th element in the original list of ages?
```{r}
my_data <- read.delim("customers.txt")# import data from customers.txt in my_data
age <- my_data[,1]                    # the array age will get the data in my_data
age[5]                                          
```
We put the data into a vector which called age and use age[5] to show the result

### 2.What is the fifth lowest age?
```{r}
sorted_age <- sort(age, decreasing = F) # sort the data in ascending order and 
sorted_age[5]                           # put the result into sorted_age
```
We use a vector to store the sorted age in ascending order and sorted_age[5] will show the fifth lowest age 

### 3.Extract the five lowest ages together
```{r}
extract_sorted_age <- sorted_age[1:5]   # put the 1st ~ 5th element of sorted_age 
extract_sorted_age                      # into extract_sorted_age
```
We use vector called extract_sorted_age to store the five lowest ages together and show the result

### 4.Get the five highest ages by first sorting them in decreasing order first.
```{r}
sorted_age <- sort(age, decreasing = TRUE)# sort the data in decreasing order and 
extract_sorted_age <- sorted_age[1:5]     # put into sorted_age and store into the
extract_sorted_age                        # extract_sorted_age
```
We use a vector to store the sorted age in decreasing order and use extract_sorted_age to store the result

### 5.What is the average (mean) age?
```{r}
mean(age)                                 # use mean() to get the average age 
```
We use mean() to get the average age

### 6.What is the standard deviation of ages?
```{r}
sd(age)                     # use sd() to get the standard deviation of age
```
We use sd() to get standard deviation of ages

### 7.Make a new variable called age_diff, with the difference between each age and the mean age
```{r}
age_diff <- age - mean(age) # let the element of age minus the average of age
age_diff                    # show the result
```
### 8.What is the average “difference between each age and the mean age”?
```{r}
mean(age_diff)
```
Since computer uses exponent notation to represent the floating number, the answer is very close but not equal to 0.




### 9.Visualize the raw data as we did in class: (a) histogram, (b) density plot, (c) boxplot+stripchart
```{r}
# (a) Plot a histogram of age
hist(age, main = "histogram of age", xlab = "age")  # draw the histogram and show the result
```
```{r}
# (b) density plot 
plot(density(age), main = "Density plot of age", xlab = "age")  # draw the density plot
```
```{r}
# (c) boxplot + stripchart
library(ggplot2)                              # use ggplot2 to draw boxplot + stripchart
ggplot(my_data, aes(x = "", y = age)) + 
  geom_boxplot()+
  geom_jitter()

```