---
title: "HW_Week15_108020033"
author: "Che-Wei, Chang"
output:
  pdf_document: default
  always_allow_html: true
date: "2023-05-25 helped by 108020024"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 1) Composite Path Models using PLS-PM

```{r}
library(seminr)
```

a. Create a PLS path model using SEMinR, with all the following characteristics:

```{r}
# Import the data
sec <- read.csv("security_data_sem.csv")
```

   (i). Measurement model – all constructs are measured as composites:
   
      1. Trust in website (TRUST): items TRST1 - TRST4
      
      2. Perceived security of website (SEC): items PSEC1 - PSEC4
      
      3. Reputation of website (REP): items PREP1 - PREP4
      
      4. Investment in website (INV): items PINV1 - PINV3
      
      5. Perception of privacy policies (POL): items PPSS1 - PPSS3
      
      6. Familiarity with website (FAML): item FAML1
         (see the documentation of SEMinR for making single item constructs)
         
      7. Interaction between REP and POL (use orthogonalized product terms)

```{r}
# Create measurement model
sec_intxn_mm <- constructs(
  composite("TRUST", multi_items("TRST", 1:4)),
  composite("SEC", multi_items("PSEC", 1:4)),
  composite("REP", multi_items("PREP", 1:4)),
  composite("INV", multi_items("PINV", 1:3)), 
  composite("POL", multi_items("PPSS", 1:3)),
  composite("FAML", single_item("FAML1")),
  interaction_term(iv = "REP", moderator = "POL", method = orthogonal)
)
```

   (ii). Structural Model – paths between constructs as shown in this causal model:
         REP + INV + POL + FAML + (REP $\times$ POL) → SEC → TRUST

```{r}
# Create structural model
sec_intxn_sm <- relationships(
  paths(from = c("REP", "INV", "POL", "FAML", "REP*POL"), to = "SEC"),
  paths(from = "SEC", to = "TRUST")
)
```

```{r}
# Run estimation algorithm 
sec_intxn_pls <- estimate_pls(
  data = sec,
  measurement_model = sec_intxn_mm,
  structural_model = sec_intxn_sm
)
```

b. Show us the following results in table or figure formats:

   (i). Plot a figure of the estimated model

```{r}
# Show the plot of the estimated model
plot(sec_intxn_pls)
```

   (ii). Weights and loadings of composites
   
```{r}
# Store the summary
sec_intxn_report <- summary(sec_intxn_pls)

# Show the weights of composites
sec_intxn_report$weights

# Show the loadings of composites
sec_intxn_report$composite_scores
```

   (iii). Regression coefficients of paths between factors
   
```{r}
# Show the regression coefficients of paths between factors
sec_intxn_report$paths
```

   (iv). Bootstrapped path coefficients: t-values, 95% CI
   
```{r}
boot_pls <- bootstrap_model(sec_intxn_pls, nboot = 1000)
summary(boot_pls)
```

### Question 2) Common-Factor Models using CB-SEM

a. Create a common factor model using SEMinR, with the following characteristics:

   (i). Either respecify all the constructs as being reflective(), or use the as.reflective() function to convert your earlier measurement model to being entirely reflective.
   
```{r}
# Create measurement model
sec_cf_mm <- constructs(
  reflective("TRUST", multi_items("TRST", 1:4)),
  reflective("SEC", multi_items("PSEC", 1:4)),
  reflective("REP", multi_items("PREP", 1:4)),
  reflective("INV", multi_items("PINV", 1:3)), 
  reflective("POL", multi_items("PPSS", 1:3)),
  reflective("FAML", single_item("FAML1")),
  interaction_term(iv = "REP", moderator = "POL", method = orthogonal)
)
```

   (ii). Use the same structural model as before (you can just reuse it again!)
   
```{r}
# Use the same structural model as before
# Run the estimation algorithm
sec_cf_pls <- estimate_cbsem(
  data = sec,
  measurement_model = sec_cf_mm,
  structural_model = sec_intxn_sm
)
```

b. Show us the following results in table or figure formats

   (i). Plot a figure of the estimated model (it will look different from your PLS model!)
   
```{r}
# Show the plot of estimated model
plot(sec_cf_pls)
```

   (ii). Loadings of composites
   
```{r}
# Store the summary
sec_cf_report <- summary(sec_cf_pls)

# Show the loadings of composites
sec_cf_report$loadings
```

   (iii). Regression coefficients of paths between factors, and their p-values

```{r}
# Show the regression coefficients of paths between factors and their p-values
sec_cf_report$paths
```
